Human Endogenous Retroviral Transcripts Show Potential as Novel, Tumor Specific Biomarkers for Colorectal Cancer

### 2021-01-30 manuscript/sups/supmethods.tex
fix tabel legends, clarify three sample changes from field effect

\subsection*{Bioinformatic processing of RNA-seq reads}
RNA-seq reads for all samples were stored and analyzed in FASTQ format.
Reads from the Genomic Data Commons were downloaded in BAM files that included unaligned reads.
All reads downloaded in BAM format were converted to FASTQ using Biobambam2 v2.0.87 \citep{Tischler2014}.

>>>>

\subsection*{Bioinformatic processing of RNA-seq reads}
RNA-seq reads for all samples were stored and analyzed in FASTQ format.
Reads from the Genomic Data Commons (GDC) were downloaded in BAM files that included unaligned reads.
All reads downloaded in BAM format were converted to FASTQ using Biobambam2 v2.0.87 \citep{Tischler2014}.

During preliminary processing of reads from the GDC, two sets of paired-end reads from tumor samples previously included in Cohort A of Dampier et al. (2020) \citep{Dampier2020} that would otherwise have been included in the discovery cohort of this study were found to be from formalin fixed, paraffin embedded as opposed to flash frozen tissue.
The two sets of reads, identified by file UUIDs b7e78979-c177-427d-a627-b5f241f36232 and ff6d6688-c19c-4a7f-8058-4d1bc0249d83 for subjects TCGA-A6-2677 and TCGA-A6-2672, respectively, were excluded from the discovery cohort.
Single-end reads from flash frozen tissue of the same tumors (file UUIDs 90832632-cf57-463b-9d08-c76975066f56 and f08dc7f4-3cc3-4743-a84e-d586d74af8d1, respectively) were added to independent cohort \#1.
With the addition of those two tumor samples to independent cohort \#1, and in order to increase the number of tumor adjacent normal samples in the discovery cohort, single-end reads from a tumor sample for subject TCGA-A6-2684 (file UUID ad1cc824-7bbc-4883-96da-55e51d1b6f0d) were removed from independent cohort \#1, and paired-end reads from tumor adjacent normal tissue from the same subject (file UUID 18625fe4-3c19-45d9-9d7c-a295fbf83f2e) were added to the discovery cohort.
Besides those changes in sample allocation, the cohorts are the same as those analyzed in Dampier et al. (2020) \citep{Dampier2020}.

### 2021-01-30 manuscript/tabs/tab_disc.tex
fix tabel legends, clarify three sample changes from field effect

\begin{table}[ht]
    \small
    \centering
    \caption{Phenotypic and demographic characteristics of discovery cohort}
    \label{tab:discovery}
    \begin{tabular}{ll|llll|lll|lll}
        \textbf{Source} & \textbf{Dataset} & \multicolumn{4}{l|}{\textbf{\# samples}} & \multicolumn{3}{l|}{\textbf{Sex (\% female)}} & \multicolumn{3}{l}{\textbf{Age (mean, sd)}} \\
        \cline{3-12}
         & & HLT & NAT & TUM & Total & HLT & NAT* & TUM* & HLT & NAT* & TUM* \\
        \hline
        BarcUVa & BarcUVa & 260 & -- & -- & 260 & 63 & -- & -- & 60, 7 & -- & -- \\
        GDC & TCGA-COAD & -- & 36 & 209 & 245 & -- & 53 & 45 & -- & 72, 13 & 65, 13 \\
        GDC & TCGA-READ & -- & 8 & 81 & 89 & -- & 88 & 46 & -- & 67, 18 & 63, 12 \\
        SRA & GTEx & 202 & -- & -- & 202 & 42 & -- & -- & 49, 13 & -- & -- \\
        SRA & HebeiMU & -- & 4 & 5 & 9 & -- & 75 & 60 & -- & \emph{NR} & \emph{NR} \\
        SRA & KoreaAMC & -- & 8 & 6 & 14 & -- & \emph{NR} & \emph{NR} & -- & \emph{NR} & \emph{NR} \\
        SRA & KoreaPNU & -- & 1 & 2 & 3 & -- & 0 & 100 & -- & 62, -- & 73, 4 \\
        SRA & Mayo & -- & 4 & 8 & 12 & -- & 50 & 38 & -- & 65, 16 & 66, 18 \\
        \hline
        Total &  & 462 & 61 & 311 & 834 & 54\% & 59\% & 46\% & 55, 11 & 70, 14 & 65, 13 \\
        \hline
        \multicolumn{12}{l}{* missing data on subset of samples; \emph{NR} signifies not reported}
    \end{tabular}
\end{table}

>>>>

\begin{table}[ht]
    \small
    \centering
    \caption{Phenotypic and demographic characteristics of discovery cohort}
    \label{tab:discovery}
    \begin{tabular}{ll|llll|lll|lll}
        \textbf{Source} & \textbf{Dataset} & \multicolumn{4}{l|}{\textbf{\# samples}} & \multicolumn{3}{l|}{\textbf{Sex (\% female)}} & \multicolumn{3}{l}{\textbf{Age (mean, sd)}} \\
        \cline{3-12}
         & & HLT & NAT & TUM & Total & HLT & NAT* & TUM* & HLT & NAT* & TUM* \\
        \hline
        BarcUVa & BarcUVa & 260 & -- & -- & 260 & 63 & -- & -- & 60, 7 & -- & -- \\
        GDC & TCGA-COAD & -- & 36 & 209 & 245 & -- & 53 & 45 & -- & 72, 13 & 65, 13 \\
        GDC & TCGA-READ & -- & 8 & 81 & 89 & -- & 88 & 46 & -- & 67, 18 & 63, 12 \\
        SRA & GTEx & 202 & -- & -- & 202 & 42 & -- & -- & 49, 13 & -- & -- \\
        SRA & HebeiMU & -- & 4 & 5 & 9 & -- & 75 & 60 & -- & \emph{NR} & \emph{NR} \\
        SRA & KoreaAMC & -- & 8 & 6 & 14 & -- & \emph{NR} & \emph{NR} & -- & \emph{NR} & \emph{NR} \\
        SRA & KoreaPNU & -- & 1 & 2 & 3 & -- & 0 & 100 & -- & 62, -- & 73, 4 \\
        SRA & Mayo & -- & 4 & 8 & 12 & -- & 50 & 38 & -- & 65, 16 & 66, 18 \\
        \hline
        Total &  & 462 & 61 & 311 & 834 & 54\% & 59\% & 46\% & 55, 11 & 70, 14 & 65, 13 \\
        \hline
        \multicolumn{12}{l}{* missing data on subset of samples; \emph{NR}: not reported; \emph{GDC}: Genomic Data Commons; \emph{SRA}: Sequence Read Archive; \emph{TCGA}: The Cancer Genome Atlas; \emph{HLT}: healthy; \emph{NAT}: normal adjacent to tumor; \emph{TUM}: tumor}
    \end{tabular}
\end{table}

### 2021-01-30 manuscript/tabs/tab_ind1.tex
fix tabel legends, clarify three sample changes from field effect

\begin{table}[ht]
    \small
    \centering
    \caption{Phenotypic and demographic characteristics of independent cohort \#1}
    \label{tab:ind1}
    \begin{tabular}{ll|llll|lll|lll}
        \textbf{Source} & \textbf{Dataset} & \multicolumn{4}{l|}{\textbf{\# samples}} & \multicolumn{3}{l|}{\textbf{Sex (\% female)}} & \multicolumn{3}{l}{\textbf{Age (mean, sd)}} \\
        \cline{3-12}
         & & HLT & NAT & TUM & Total & HLT & NAT* & TUM* & HLT & NAT* & TUM* \\
        \hline
        GDC & TCGA-COAD & -- & -- & 135 & 135 & -- & -- & 47 & -- & -- & 69, 12 \\
        GDC & TCGA-READ & -- & -- & 60 & 60 & -- & -- & 42 & -- & -- & 68, 10 \\
        SRA & MtSinai & -- & 47 & -- & 47 & -- & 62 & -- & -- & 65, 13 & -- \\
        SRA & CityOfHope & -- & 2 & 8 & 10 & -- & \emph{NR} & \emph{NR} & -- & \emph{NR} & \emph{NR} \\
        SRA & Singapore & -- & 5 & 2 & 7 & -- & \emph{NR} & \emph{NR} & -- & \emph{NR} & \emph{NR} \\
        SRA & Utah & 10 & -- & -- & 10 & 40 & -- & -- & 55, 7 & -- & -- \\
        SRA & ZhejiangU & -- & 2 & 2 & 4 & -- & 0 & 0 & -- & 59, 30 & 54, 2 \\
        SRA & TexasAM & 2 & -- & -- & 2 & 50 & -- & -- & 39, 8 & -- & -- \\
        \hline
        Total &  & 12 & 56 & 207 & 275 & 42\% & 59\% & 45\% & 52, 10 & 65, 14 & 69, 11 \\
        \hline
        \multicolumn{12}{l}{* missing data on subset of samples; \emph{NR} signifies not reported} \\
    \end{tabular}
\end{table}

>>>>

\begin{table}[ht]
    \small
    \centering
    \caption{Phenotypic and demographic characteristics of independent cohort \#1}
    \label{tab:ind1}
    \begin{tabular}{ll|llll|lll|lll}
        \textbf{Source} & \textbf{Dataset} & \multicolumn{4}{l|}{\textbf{\# samples}} & \multicolumn{3}{l|}{\textbf{Sex (\% female)}} & \multicolumn{3}{l}{\textbf{Age (mean, sd)}} \\
        \cline{3-12}
         & & HLT & NAT & TUM & Total & HLT & NAT* & TUM* & HLT & NAT* & TUM* \\
        \hline
        GDC & TCGA-COAD & -- & -- & 135 & 135 & -- & -- & 47 & -- & -- & 69, 12 \\
        GDC & TCGA-READ & -- & -- & 60 & 60 & -- & -- & 42 & -- & -- & 68, 10 \\
        SRA & MtSinai & -- & 47 & -- & 47 & -- & 62 & -- & -- & 65, 13 & -- \\
        SRA & CityOfHope & -- & 2 & 8 & 10 & -- & \emph{NR} & \emph{NR} & -- & \emph{NR} & \emph{NR} \\
        SRA & Singapore & -- & 5 & 2 & 7 & -- & \emph{NR} & \emph{NR} & -- & \emph{NR} & \emph{NR} \\
        SRA & Utah & 10 & -- & -- & 10 & 40 & -- & -- & 55, 7 & -- & -- \\
        SRA & ZhejiangU & -- & 2 & 2 & 4 & -- & 0 & 0 & -- & 59, 30 & 54, 2 \\
        SRA & TexasAM & 2 & -- & -- & 2 & 50 & -- & -- & 39, 8 & -- & -- \\
        \hline
        Total &  & 12 & 56 & 207 & 275 & 42\% & 59\% & 45\% & 52, 10 & 65, 14 & 69, 11 \\
        \hline
        \multicolumn{12}{l}{* missing data on subset of samples; \emph{NR}: not reported; \emph{GDC}: Genomic Data Commons; \emph{SRA}: Sequence Read Archive; \emph{TCGA}: The Cancer Genome Atlas; \emph{HLT}: healthy; \emph{NAT}: normal adjacent to tumor; \emph{TUM}: tumor} \\
    \end{tabular}
\end{table}

### 2021-01-30 manuscript/tabs/tab_ind2.tex
fix tabel legends, clarify three sample changes from field effect

\begin{table}[ht]
    \small
    \centering
    \caption{Phenotypic and demographic characteristics of independent cohort \#2}
    \label{tab:ind2}
    \begin{tabular}{ll|llll|lll|lll}
        \textbf{Source} & \textbf{Dataset} & \multicolumn{4}{l|}{\textbf{\# samples}} & \multicolumn{3}{l|}{\textbf{Sex (\% female)}} & \multicolumn{3}{l}{\textbf{Age (mean, sd)}} \\
        \cline{3-12}
         & & HLT & NAT & TUM & Total & HLT & NAT* & TUM* & HLT & NAT* & TUM* \\
        \hline
        GDC & TCGA-COAD & -- & 3 & 3 & 6 & -- & 50 & 50 & -- & 62, 13 & 62, 13 \\
        GDC & TCGA-READ & -- & 1 & 1 & 2 & -- & 0 & 0 & -- & 50, -- & 50, -- \\
        SRA & HebeiMU & -- & 1 & 1 & 2 & -- & 0 & 0 & -- & \emph{NR} & \emph{NR} \\
        SRA & KoreaAMC & -- & 4 & 4 & 8 & -- & \emph{NR} & \emph{NR} & -- & \emph{NR} & \emph{NR} \\
        SRA & KoreaPNU & -- & 2 & 2 & 4 & -- & 50 & 50 & -- & 74, 1 & 74, 1 \\
        SRA & Mayo & -- & 4 & 4 & 8 & -- & 50 & 50 & -- & 70, 10 & 70, 10 \\
        \hline
        Total &  & -- & 15 & 15 & 30 & -- & 36\% & 36\% & -- & 67, 12 & 67, 12 \\
        \hline
        \multicolumn{12}{l}{* missing data on subset of samples; \emph{NR} signifies not reported} \\
    \end{tabular}
\end{table}

>>>>

\begin{table}[ht]
    \small
    \centering
    \caption{Phenotypic and demographic characteristics of independent cohort \#2}
    \label{tab:ind2}
    \begin{tabular}{ll|llll|lll|lll}
        \textbf{Source} & \textbf{Dataset} & \multicolumn{4}{l|}{\textbf{\# samples}} & \multicolumn{3}{l|}{\textbf{Sex (\% female)}} & \multicolumn{3}{l}{\textbf{Age (mean, sd)}} \\
        \cline{3-12}
         & & HLT & NAT & TUM & Total & HLT & NAT* & TUM* & HLT & NAT* & TUM* \\
        \hline
        GDC & TCGA-COAD & -- & 3 & 3 & 6 & -- & 50 & 50 & -- & 62, 13 & 62, 13 \\
        GDC & TCGA-READ & -- & 1 & 1 & 2 & -- & 0 & 0 & -- & 50, -- & 50, -- \\
        SRA & HebeiMU & -- & 1 & 1 & 2 & -- & 0 & 0 & -- & \emph{NR} & \emph{NR} \\
        SRA & KoreaAMC & -- & 4 & 4 & 8 & -- & \emph{NR} & \emph{NR} & -- & \emph{NR} & \emph{NR} \\
        SRA & KoreaPNU & -- & 2 & 2 & 4 & -- & 50 & 50 & -- & 74, 1 & 74, 1 \\
        SRA & Mayo & -- & 4 & 4 & 8 & -- & 50 & 50 & -- & 70, 10 & 70, 10 \\
        \hline
        Total &  & -- & 15 & 15 & 30 & -- & 36\% & 36\% & -- & 67, 12 & 67, 12 \\
        \hline
        \multicolumn{12}{l}{* missing data on subset of samples; \emph{NR}: not reported; \emph{GDC}: Genomic Data Commons; \emph{SRA}: Sequence Read Archive; \emph{TCGA}: The Cancer Genome Atlas; \emph{HLT}: healthy; \emph{NAT}: normal adjacent to tumor; \emph{TUM}: tumor} \\
    \end{tabular}
\end{table}
